---
title: Draw a handball court with ggplot
author: Valery Zúñiga
date: '2021-01-09'
slug: draw-a-handball-court-with-ggplot
categories:
  - R
  - handball
  - visualization
tags:
  - ggplot
  - handball
  - rstats
---


The strategy to draw a court is to define the different lines and use geom_path to draw each segment.
For circles, we use the following function to create the x-y coordinates. 
For lines, we just define x-y start and end coordinates.

```{r}
library(ggplot2)

draw_circle <- function(center = c(0, 0),
                        diameter = 1, 
                        npoints = 12000, 
                        start = 0, 
                        end = 2){
  
  tt <- seq(start*pi, end*pi, length.out = npoints)
  
  data.frame(
    x = center[1] + diameter / 2 * cos(tt),
    y = center[2] + diameter / 2 * sin(tt)
  )

}

```

Now, we can define each line of the court. 
Our court dimensions go from -20 to 20 by -10 to 10, which would be 40x20. 


```{r}

side_and_goal_lines <- 
  data.frame(x = c(-20, 20, 20, -20, -20),
             y = c(-10, -10, 10, 10, -10))

center_line <- 
  data.frame(x = c(0, 0),
             y = c(-10, 10))

goal1 <- 
  data.frame(x = c(-20, -20),
             y = c(-1.5, 1.5))

goal2 <- 
  data.frame(x = c(20, 20),
             y = c(-1.5, 1.5))


front6_1 <- 
  data.frame(x = c(-14, -14),
             y = c(-1.5, 1.5))

front6_2 <- 
  data.frame(x = c(14, 14),
             y = c(-1.5, 1.5))


quart_circle6_1_1 <- 
  draw_circle(center = c(-20, -1.5), 
              diameter = 12, 
              start = 1.5, end = 2)

quart_circle6_1_2 <- 
  draw_circle(center = c(-20, 1.5), 
              diameter = 12, 
              start = 0, end = 0.5)

quart_circle6_2_1 <- 
  draw_circle(center = c(20, -1.5), 
              diameter = 12, 
              start = 1, end = 1.5)

quart_circle6_2_2 <- 
  draw_circle(center = c(20, 1.5), 
              diameter = 12,
              start = 0.5, end = 1)

# line 
free_thr9_1 <- 
  data.frame(x = c(-11, -11),
             y = c(-1.5, 1.5))
# line 
free_thr9_2 <- 
  data.frame(x = c(11, 11),
             y = c(-1.5, 1.5))


free_thr9_1_1 <- 
  draw_circle(center = c(-20, -1.5),
              diameter = 18, 
              start = 1.62, end = 2)

free_thr9_1_2 <- 
  draw_circle(center = c(-20, 1.5),
              diameter = 18, 
              start = 0, end = 0.38)


free_thr9_2_1 <- 
  draw_circle(center = c(20, -1.5),
              diameter = 18, 
              start = 1, end = 1.38)

free_thr9_2_2 <- 
  draw_circle(center = c(20, 1.5), 
              diameter = 18, 
              start = 0.62, end = 1)


goal_restr_1 <-
  data.frame(x = c(-16, -16),
             y = c(-0.075, 0.075))

goal_restr_2 <-
  data.frame(x = c(16, 16),
             y = c(-0.075, 0.075))

sevenm_1 <- 
  data.frame(x = c(-13, -13),
             y = c(-0.5, 0.5))

sevenm_2 <-
  data.frame(x = c(13, 13),
             y = c(-0.5, 0.5))

substitution_1 <- 
  data.frame(x = c(-4.5, -4.5),
             y = c(-10.15, -9.85))

substitution_2 <- 
  data.frame(x = c(4.5, 4.5),
             y = c(-10.15, -9.85))

side_and_goal_lines_half <- 
  data.frame(x = c(5, 20, 20, 5, 5),
             y = c(-10, -10, 10, 10, -10))


```

We are ready to pass these lines to a ggplot for a complete court.
Let's create a function so we can have some parameters for our court and then ease to call our plot.

As parameters:
- $vertical$: If we want to see a vertical plot
- $flip$: Flips the plot 180°
- $plotly$: Indicates if we want a plotly instead of a classic ggplot. 

```{r}
court <- function(vertical = FALSE, flip = FALSE, plotly = FALSE){

  make_flip <- NULL
  
  if(flip) make_flip <- '-'
  
  type_of_plot <- 
    ifelse(vertical,
           paste0('aes(y, ', make_flip, 'x)'),
           paste0('aes(', make_flip, 'x, ', 'y)'))
  
  
  plot <- ggplot(mapping = eval(parse(text = type_of_plot))) +
    geom_path(data = side_and_goal_lines, size = 1) +
    geom_path(data = center_line, size = 1) +
    geom_path(data = goal1, color = 'red', size = 1.5) +
    geom_path(data = goal2, color = 'red', size = 1.5) +
    geom_path(data = front6_1, size = 1) +
    geom_path(data = front6_2, size = 1) +
    geom_path(data = quart_circle6_1_1, size = 1) +
    geom_path(data = quart_circle6_1_2, size = 1) +
    geom_path(data = quart_circle6_2_1, size = 1) +
    geom_path(data = quart_circle6_2_2, size = 1) +
    geom_path(data = free_thr9_1, linetype = 'dashed', size = 1) +
    geom_path(data = free_thr9_2, linetype = 'dashed', size = 1) +
    geom_path(data = free_thr9_1_1, linetype = 'dashed', size = 1) +
    geom_path(data = free_thr9_1_2, linetype = 'dashed', size = 1) +
    geom_path(data = free_thr9_2_1, linetype = 'dashed', size = 1) +
    geom_path(data = free_thr9_2_2, linetype = 'dashed', size = 1) +
    geom_path(data = goal_restr_1, size = 1) +
    geom_path(data = goal_restr_2, size = 1) +
    geom_path(data = sevenm_1, size = 1) +
    geom_path(data = sevenm_2, size = 1) +
    geom_path(data = substitution_1, size = 1) +
    geom_path(data = substitution_2, size = 1) +
    coord_fixed() + # We want to maintain the 40x20 proportion 
    theme_void()
  
  if(plotly) plot <- plotly::ggplotly(plot)
  
  return(plot)
}

```

And half a court with the same parameters. 

```{r}
half_court <- function(vertical = TRUE, flip = FALSE, plotly = FALSE){
  
  make_flip <- NULL
  
  if(flip) make_flip <- '-'
  
  type_of_plot <- 
    ifelse(vertical,
           paste0('aes(y, ', make_flip, 'x)'),
           paste0('aes(', make_flip, 'x, ', 'y)'))
  
  
  plot <- ggplot(mapping = eval(parse(text = type_of_plot))) +
    geom_path(data = side_and_goal_lines_half, size = 1) +
    geom_path(data = goal2, color = 'red', size = 1.5) +
    geom_path(data = front6_2, size = 1) +
    geom_path(data = quart_circle6_2_1, size = 1) +
    geom_path(data = quart_circle6_2_2, size = 1) +
    geom_path(data = free_thr9_2, linetype = 'dashed', size = 1) +
    geom_path(data = free_thr9_2_1, linetype = 'dashed', size = 1) +
    geom_path(data = free_thr9_2_2, linetype = 'dashed', size = 1) +
    geom_path(data = goal_restr_2, size = 1) +
    geom_path(data = sevenm_2, size = 1) +
    coord_fixed() +
    theme_void()
  
  if(plotly) plot <- plotly::ggplotly(plot)
  
  return(plot)
}
```

No we draw a court:
```{r}
court(plotly = TRUE)

```

And half a court: 


```{r}
half_court(vertical = TRUE, flip = TRUE, plotly = FALSE)
```

As this object ended with a theme_void() call at the end of our function, we can extend this plot and add other information on top. 

For example, let's generate some data on where some shots were taken and if they were a goal or not:

```{r}
shots <-
  data.frame(x = c(-13, -12, 11, -11, 9.5),
             y = c(2, 5, -3, -1, 0),
             goal = c(1, 0, 1, 1, 0))
```


Now we just generate out court and then add new layers.
```{r}
campo() +
      geom_point(data = shots, aes(x, y), color = ifelse(shots$goal == 1, 'Green', 'Red'), size = 6)
```






